---
title: "spacetimeadultanalysis"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r load-libraries, library, message=F, results='hide', warning=FALSE}
library(knitr)
library(tidyboot)
library(tidyverse)
library(here)

theme_set(theme_classic(base_size = 16))
```

```{r message = FALSE}
# a function to get mean exponentiated distances for word types
get_means <- function(data) {
  newdata <- data %>%
    group_by(spacetime, word) %>%
    summarise_at(vars(distance,expdist,inversedist), mean) %>%
    tidyboot_mean(expdist) %>%
    mutate(spacetime = factor(spacetime, levels = c("other", "both", "space", "time"))) 
  return(newdata)
}
```

```{r message = FALSE}
# a function to get summed inverse exponentiated distances for word types
get_sums <- function(data) {
  newdata <- data %>%
    group_by(spacetime, word) %>%
    summarise_at(vars(inversedist), sum) %>%
    tidyboot(inversedist, summary_function = sum(inversedist), 
             statistics_functions =  list("ci_lower" = ci_lower,
                                          "sum" = mean,
                                          "ci_upper" = ci_upper))
  return(newdata)
}
```

```{r message = FALSE}
# a function to get mean exponentiated distances for word types
read_data <- function(data) {
  cleandata <- data %>%
  mutate(spacetime = ifelse(spacetime == "O", "other",spacetime),
         spacetime = ifelse(spacetime == "S", "space",spacetime),
         spacetime = ifelse(spacetime == "T", "time",spacetime),
         spacetime = ifelse(spacetime == "B", "both",spacetime),
         spacetime = as_factor(spacetime)) %>%
  mutate(spacetime = as_factor(spacetime)) %>%
  mutate(expdist = exp(distance)) %>%
  mutate(inversedist = 1/expdist) %>%
  mutate(distbin = cut(expdist, breaks = 10, labels = FALSE))
  return(cleandata)
}
```

```{r message = FALSE, warning=FALSE}
longdata <- read_data(read_csv(here("neighbors/cocalongneighbors.csv")))

shortdata <- read_data(read_csv(here("neighbors/cocashortneighbors.csv"))) 

widedata <- read_data(read_csv(here("neighbors/cocawideneighbors.csv"))) 

soondata <- read_data(read_csv(here("neighbors/cocasoonneighbors.csv"))) 

longmeans <- get_means(longdata)
shortmeans <- get_means(shortdata)
soonmeans <- get_means(soondata)
widemeans <- get_means(widedata)

longsums <- get_sums(longdata)
shortsums <- get_sums(shortdata)
soonsums <- get_sums(soondata)
widesums <- get_sums(widedata)

```

Below: A histogram of exponentiated word distances from "long". These are the closest 500 words.

Exponentiating the distances spaces out further distances. This helps deal with the fact that in any space, there are more things further from you than closer to you: a unit increase in radius corresponds to a unit squared increase in area covered. The transformation helps normalize the distribution of words over distance. Throughout, we will be using exponentiated distances.

```{r}
hist(longdata$expdist)
```



Below: A histogram of exponentiated word distances from "short". These are the closest 500 words.

```{r}
hist(shortdata$expdist)
```



Below: A histogram of exponentiated word distances from "soon". These are the closest 100 words.

```{r}
hist(soondata$expdist)
```



Below: A histogram of exponentiated word distances from "wide". These are the closest 100 words.

```{r}
hist(widedata$expdist)
```



Below: Mean exponentiated distances of space, time, both and other words from "long". Error bars here and throughout are 95% bootstrapped confidence intervals.

```{r}
longmeans %>%
  ggplot(aes(x = spacetime, y = empirical_stat)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  ggtitle("Average exponentiated distances from long")
```

Below: Summed exponentiated distances of space, time, both and other words from "long". Error bars here and throughout are 95% bootstrapped confidence intervals.

```{r}
longsums %>%
  ggplot(aes(x = spacetime, y = sum)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  ggtitle("Summed inverse exponentiated distances from long")
```



Below: Mean exponentiated distances of space, time, both and other words from "short".

```{r}
shortmeans %>%
  ggplot(aes(x = spacetime, y = empirical_stat)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  ggtitle("Average exponentiated distances from short")
```

Below: Summed inverse exponentiateddistances of space, time, both and other words from "short".

```{r}
shortsums %>%
  ggplot(aes(x = spacetime, y = sum)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  ggtitle("Summed inverse exponentiated distances from short")
```



Below: Mean exponentiated distances of space, time, both and other words from "soon". Note: there were no pure space words for 'soon'.

```{r}
soonmeans %>%
  ggplot(aes(x = spacetime, y = empirical_stat)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  ggtitle("Average exponentiated distances from soon") +
  geom_text(aes(label = n), vjust=-4, size = 6)
```

Below: Summed exponentiated distances of space, time, both and other words from "soon". Note: there were no pure space words for 'soon'.

```{r}
soonsums %>%
  ggplot(aes(x = spacetime, y = sum)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  ggtitle("Summed inverse exponentiated distances from soon") +
  geom_text(aes(label = n), vjust=-4, size = 6)
```



Below: Mean exponentiated distances of space, time, both and other words from "wide". Note: there were no pure time words for 'wide'.

```{r}
widemeans %>%
  ggplot(aes(x = spacetime, y = empirical_stat)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  ggtitle("Average exponentiated distances from wide") +
  geom_text(aes(label = n), vjust=-4, size = 6)
```

Below: Summed exponentiated distances of space, time, both and other words from "wide". Note: there were no pure time words for 'wide'.

```{r}
widesums %>%
  ggplot(aes(x = spacetime, y = sum)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  ggtitle("Summed inverse exponentiated distances from wide") +
  geom_text(aes(label = n), vjust=-4, size = 6)
```

Below: A plot of mean distances of space, time, both and other words from "long" within the closest 20, 100, and 500 neighbors. The numbers on the plot denote the number of words of that type in that range of neighbors.

```{r}
longmeans$threshold = "closest 500"

longmeans20 <- longdata %>%
  top_n(20, -expdist) %>%
  get_means() %>%
  mutate(threshold = "closest 20")

longmeans100 <- longdata %>%
  top_n(100, -expdist) %>%
  get_means() %>%
  mutate(threshold = "closest 100")

longmeansAll <- longmeans100 %>%
  rbind(longmeans20) %>%
  rbind(longmeans) %>%
  mutate(threshold = factor(threshold, levels = c("closest 20", "closest 100", "closest 500"))) %>%
  mutate(spacetime = factor(spacetime, levels = c("other", "both", "space", "time")))

longmeansAll %>%
  ggplot(aes(x = spacetime, y = empirical_stat)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~threshold) +
  geom_text(aes(label = n), vjust=-4, size = 6)
```


Below: A plot of mean distances of space, time, both and other words from "short" within the closest 20, 100, and 500 neighbors.

```{r}
shortmeans$threshold = "closest 500"

shortmeans20 <- shortdata %>%
  top_n(20, -expdist) %>%
  get_means() %>%
  mutate(threshold = "closest 20")

shortmeans100 <- shortdata %>%
  top_n(100, -expdist) %>%
  get_means() %>%
  mutate(threshold = "closest 100")

shortmeansAll <- shortmeans100 %>%
  rbind(shortmeans20) %>%
  rbind(shortmeans) %>%
  mutate(threshold = factor(threshold, levels = c("closest 20", "closest 100", "closest 500"))) %>%
  mutate(spacetime = factor(spacetime, levels = c("other", "both", "space", "time"))) 

shortmeansAll %>%
  ggplot(aes(x = spacetime, y = empirical_stat)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~threshold) +
  geom_text(aes(label = n), vjust=-4, size = 6) 

```



Below: Linear model predicting distance from "long" by word type (space, time, etc.) in the closest 100 words, where we're more likely to see differences. Space is the reference category. Space and time are not significantly different.

```{r}
longmodeldata <- longdata %>%
  top_n(100, -expdist) %>%
  mutate(spacetime = relevel(spacetime, ref = 'space'))
         
longmodel <- glm(expdist ~ spacetime, family = gaussian, data = longmodeldata)

summary(longmodel)
```



Below: Linear model predicting distance from "short" by word type (space, time, etc.) in the closest 100 words, where we're more likely to see differences. Space is the reference category. Space and time are not significantly different.

```{r}
shortmodeldata <- shortdata %>%
  top_n(100, -expdist) %>%
  mutate(spacetime = relevel(spacetime, ref = 'space'))
         
shortmodel <- glm(expdist ~ spacetime, family = gaussian, data = shortmodeldata)

summary(shortmodel)
```



Below: Linear model predicting distance from "soon" by word type (space, time, etc.) in the closest 100 words. Time is significantly different from 'both' and 'other'; there's only one space word, so the difference in distance probably can't be reliably predicted.

```{r}
soonmodeldata <- soondata %>%
  top_n(100, -expdist) %>%
  mutate(spacetime = relevel(spacetime, ref = 'time'))
         
soonmodel <- glm(expdist ~ spacetime, family = gaussian, data = soonmodeldata)

summary(soonmodel)
```



Below: Linear model predicting distance from "wide" by word type (space, time, etc.) in the closest 100 words. Space is significantly different from 'other', marginally from 'both'; there are no time words to predict from.

```{r}
widemodeldata <- widedata %>%
  top_n(100, -expdist) %>%
  mutate(spacetime = relevel(spacetime, ref = 'space'))
         
widemodel <- glm(expdist ~ spacetime, family = gaussian, data = widemodeldata)

summary(widemodel)
```