---
title: "spacetimechildesanalysis"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r load-libraries, library, message=F, results='hide', warning=FALSE}
library(knitr)
library(tidyboot)
library(tidyverse)
library(here)

theme_set(theme_classic(base_size = 16))
```


```{r message = FALSE}
read_data <- function(data) {
  cleandata <- data %>%
    mutate(spacetime = ifelse(spacetime == "O", "other",spacetime),
         spacetime = ifelse(spacetime == "S", "space",spacetime),
         spacetime = ifelse(spacetime == "T", "time",spacetime),
         spacetime = ifelse(spacetime == "B", "both",spacetime),
         spacetime = as_factor(spacetime)) %>%
    mutate(spacetime = as_factor(spacetime)) 
  return(cleandata)
}
```

```{r message = FALSE, warning=FALSE}
longdata <- read_data(read_csv(here("neighbors/long_childes_coded.csv")))

shortdata <- read_data(read_csv(here("neighbors/short_childes_coded.csv"))) 

widedata <- read_data(read_csv(here("neighbors/wide_childes_coded.csv"))) 

soondata <- read_data(read_csv(here("neighbors/soon_childes_coded.csv"))) 
```


```{r}
getSpace <- function(data){
  return(as.numeric(data$n[data$spacetime == "space"]))
}

getTime <- function(data){
  return(as.numeric(data$n[data$spacetime == "time"]))
}

getDiff <- function(data){
  return(as.numeric(data$n[data$spacetime == "time"] - data$n[data$spacetime == "space"]))
}

```


Below: Bootstrapping (n=100) samples of number of space words, time words, and their difference from the sample of 6000 random words.

Import the random sample from the corpus with 3000 words. 

```{r message = FALSE}
null_data <- read_csv(here("neighbors/random_childes_coded.csv")) %>%
    mutate(spacetime = ifelse(is.na(spacetime), "other", spacetime)) %>%
    mutate(spacetime = as_factor(spacetime)) 
```


Generate the null distribution for number of space words, time words, and their difference. 

```{r}
set.seed(185)

sampleData <- as.data.frame(rep(0, 100))

simulate_null <- function(getX) {
    sampleData <- mutate(sampleData, spacetime = sample(null_data$spacetime, size = 100, replace=TRUE))
    sampledValues <- sampleData %>%
      group_by(spacetime, .drop = FALSE) %>%
      count() 
  return(getX(sampledValues))
}

nullSpace <- replicate(10000, simulate_null(getSpace)) 
nullTime <- replicate(10000, simulate_null(getTime))  
nullDiff <- replicate(10000, simulate_null(getDiff)) 

nullSpaceDF <- data.frame(nullSpace)
nullTimeDF <- data.frame(nullTime)
nullDiffDF <- data.frame(nullDiff)
```


Below: Same analyses with the 500 observed sample and bootstrapped samples with N=500


Below: Empirical values for the number of space words, the number of time words and the difference between the two within the first 500 nearest neigbors of long.  

```{r}
empiricalValues500 <- longdata %>%
  group_by(spacetime) %>%
  count()

observedSpaceLong500 <- getSpace(empiricalValues500)
observedTimeLong500 <- getTime(empiricalValues500)
observedDiffLong500 <- getDiff(empiricalValues500)
```

Below: Empirical values for the number of space words, the number of time words and the difference between the two within the first 500 nearest neigbors of short  

```{r}
empiricalValuesShort500 <- shortdata %>%
  group_by(spacetime) %>%
  count()

observedSpaceShort500 <- getSpace(empiricalValuesShort500)
observedTimeShort500 <- getTime(empiricalValuesShort500)
observedDiffShort500 <- getDiff(empiricalValuesShort500)
```


Generate the null distribution for number of space words, time words, and their difference -- samples with n = 500

```{r}
set.seed(188)

sampleData <- as.data.frame(rep(0, 500))

simulate_null <- function(getX) {
    sampleData <- mutate(sampleData, spacetime = sample(null_data$spacetime, size = 500, replace=TRUE))
    sampledValues <- sampleData %>%
      group_by(spacetime, .drop = FALSE) %>%
      count() 
  return(getX(sampledValues))
}

nullSpace500 <- replicate(10000, simulate_null(getSpace)) 
nullTime500 <- replicate(10000, simulate_null(getTime))  
nullDiff500 <- replicate(10000, simulate_null(getDiff)) 

nullSpaceDF500 <- data.frame(nullSpace500)
nullTimeDF500 <- data.frame(nullTime500)
nullDiffDF500 <- data.frame(nullDiff500)
```




```{r}
ggplot(nullTimeDF500, aes(x=nullTime500)) +
  geom_histogram() +
  geom_vline(xintercept = observedTimeLong500, linetype="dotted", 
                color = "blue", size=1.5) +
  geom_vline(xintercept = observedTimeShort500, linetype="dotted", 
                color = "green", size=1.5) +
  theme_classic() +
  #ggtitle("Observed value of number of ") + 
  geom_text(aes(x=observedTimeLong500, label="observed count \nfor long", y=4000), colour="blue", text=element_text(size=11))+
  geom_text(aes(x=observedTimeShort500, label="observed count \nfor short", y=4000), colour="green", text=element_text(size=11)) +
  ggtitle("Null distribution for the number of time words in samples of 500") +
  labs(x = "Number of time words (out of 500)", y = "Number of samples") 
  
```


```{r}
ggplot(nullSpaceDF500, aes(x=nullSpace500)) +
  geom_histogram() +
  geom_vline(xintercept = observedSpaceLong500, linetype="dotted", 
                color = "blue", size=1.5) +
  geom_vline(xintercept = observedSpaceShort500, linetype="dotted", 
                color = "green", size=1.5) +
  theme_classic() +
  labs(x = "Number of space words", y = "Count among 10000 samples") +
  #ggtitle("Observed value of number of ") + 
  geom_text(aes(x=observedSpaceLong500, label="observed count \nfor long", y=4000), colour="blue", text=element_text(size=11)) +
  geom_text(aes(x=observedSpaceLong500, label="observed count \nfor short", y=4000), colour="green", text=element_text(size=11)) +
  ggtitle("Null distribution for the number of space words in samples of 500") +
  labs(x = "Number of space words (out of 500)", y = "Number of samples") 

```

```{r}
ggplot(nullDiffDF500, aes(x=nullDiff500)) +
  geom_histogram() +
  geom_vline(xintercept = observedDiffLong500, linetype="dotted", 
                color = "blue", size=1.5) +
  geom_vline(xintercept = observedDiffShort500, linetype="dotted", 
                color = "green", size=1.5) +
 theme_classic() +
  #labs(x = "Number of time words minus number of space words", y = "Count among 10000 samples") +
  #ggtitle("Observed value of number of ") + 
  geom_text(aes(x=observedDiffLong500, label="observed count \nfor long", y=4000), colour="blue", text=element_text(size=20)) +
  geom_text(aes(x=observedDiffShort500, label="observed count \nfor short", y=4000), colour="green", text=element_text(size=20)) +
  theme(axis.text=element_text(size=14),
          axis.title=element_text(size=16))  +
  ggtitle("Null distribution for the difference in the number of time words and space words \nin samples of 500") +
  labs(x = "Number of time - number of space words (out of 500)", y = "Number of samples") 

```

Stats for long with samples N=500

For long

```{r}
mean(nullDiffDF500 >= observedDiffLong500)
mean(nullTimeDF500 >= observedTimeLong500)
mean(nullSpaceDF500 >= observedSpaceLong500)
```

For short

```{r}
mean(nullDiffDF500 >= observedDiffShort500)
mean(nullTimeDF500 >= observedTimeShort500)
mean(nullSpaceDF500 >= observedSpaceShort500)
```


