---
title: "spacetimeadultanalysis"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r load-libraries, library, message=F, results='hide', warning=FALSE}
library(knitr)
library(tidyboot)
library(tidyverse)
library(here)

theme_set(theme_classic(base_size = 16))
```

```{r message = FALSE}
# a function to get mean similarities for word types
get_means <- function(data) {
  newdata <- data %>%
    group_by(spacetime, word) %>%
    summarise_at(vars(similarity), mean) %>%
    tidyboot_mean(similarity) %>%
    mutate(spacetime = factor(spacetime, levels = c("other", "both", "space", "time"))) 
  return(newdata)
}
```

```{r message = FALSE}
# a function to get summed similarities for word types
get_sums <- function(data) {
  newdata <- data %>%
    group_by(spacetime, word) %>%
    summarise_at(vars(similarity), sum) %>%
    tidyboot(similarity, summary_function = sum,
             statistics_functions =  list("ci_lower" = ci_lower,
                                          "ci_upper" = ci_upper))
  return(newdata)
}
```


```{r message = FALSE}
read_data <- function(data) {
  cleandata <- data %>%
    mutate(spacetime = ifelse(spacetime == "O", "other",spacetime),
         spacetime = ifelse(spacetime == "S", "space",spacetime),
         spacetime = ifelse(spacetime == "T", "time",spacetime),
         spacetime = ifelse(spacetime == "B", "both",spacetime),
         spacetime = as_factor(spacetime)) %>%
    mutate(spacetime = as_factor(spacetime)) %>%
    mutate(similarity = as.numeric(similarity)) 
  return(cleandata)
}
```

```{r message = FALSE, warning=FALSE}
longdata <- read_data(read_csv(here("neighbors/cocalongneighbors.csv")))

shortdata <- read_data(read_csv(here("neighbors/cocashortneighbors.csv"))) 

widedata <- read_data(read_csv(here("neighbors/cocawideneighbors.csv"))) 

soondata <- read_data(read_csv(here("neighbors/cocasoonneighbors.csv"))) 
```



```{r}
longmeans <- get_means(longdata)
shortmeans <- get_means(shortdata)
soonmeans <- get_means(soondata)
widemeans <- get_means(widedata)

longsums <- get_sums(longdata)
shortsums <- get_sums(shortdata)
soonsums <- get_sums(soondata)
widesums <- get_sums(widedata)
```

Below: A histogram of word similarities to "long".

```{r}
hist(longdata$similarity)
```



Below: A histogram of word similarities to "short".

```{r}
hist(shortdata$similarity)
```



Below: A histogram of word similarities to "soon".

```{r}
hist(soondata$similarity)
```



Below: A histogram of word similarities to "wide".

```{r}
hist(widedata$similarity)
```



Below: Mean similarities of space, time, both and other words to "long". Error bars here and throughout are 95% bootstrapped confidence intervals.

```{r}
longmeans %>%
  ggplot(aes(x = spacetime, y = empirical_stat)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  ggtitle("Average similarities to long")
```

Below: Summed similarities of space, time, both and other words to "long". Error bars here and throughout are 95% bootstrapped confidence intervals.

```{r}
longsums %>%
  ggplot(aes(x = spacetime, y = empirical_sum)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  ggtitle("Summed similarities to long")
```



Below: Mean similarities of space, time, both and other words to "short".

```{r}
shortmeans %>%
  ggplot(aes(x = spacetime, y = empirical_stat)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  ggtitle("Average similarities to short")
```

Below: Summed similarities of space, time, both and other words to "short".

```{r}
shortsums %>%
  ggplot(aes(x = spacetime, y = empirical_sum)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  ggtitle("Summed similarities to short")
```



Below: Mean similarities of space, time, both and other words to "soon". Note: there were no pure space words for 'soon'.

```{r}
soonmeans %>%
  ggplot(aes(x = spacetime, y = empirical_stat)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  ggtitle("Average similarities to soon") +
  geom_text(aes(label = n), vjust=-4, size = 6)
```

Below: Summed similarities of space, time, both and other words to "soon". Note: there were no pure space words for 'soon'.

```{r}
soonsums %>%
  ggplot(aes(x = spacetime, y = empirical_sum)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  ggtitle("Summed similarities to soon") +
  geom_text(aes(label = n), vjust=-4, size = 6)
```



Below: Mean similarities of space, time, both and other words to "wide". Note: there were no pure time words for 'wide'.

```{r}
widemeans %>%
  ggplot(aes(x = spacetime, y = empirical_stat)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  ggtitle("Average similarities to wide") +
  geom_text(aes(label = n), vjust=-4, size = 6)
```

Below: Summed similarities of space, time, both and other words to "wide". Note: there were no pure time words for 'wide'.

```{r}
widesums %>%
  ggplot(aes(x = spacetime, y = empirical_sum)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  ggtitle("Summed similarities to wide") +
  geom_text(aes(label = n), vjust=-4, size = 6)
```

Below: A plot of mean similarities of space, time, both and other words to "long" within the closest 20, 100, and 500 neighbors. The numbers on the plot denote the number of words of that type in that range of neighbors.

```{r}
longmeans$threshold = "closest 500"

longmeans20 <- longdata %>%
  top_n(20, similarity) %>%
  get_means() %>%
  mutate(threshold = "closest 20")

longmeans100 <- longdata %>%
  top_n(100, similarity) %>%
  get_means() %>%
  mutate(threshold = "closest 100")

longmeansAll <- longmeans100 %>%
  rbind(longmeans20) %>%
  rbind(longmeans) %>%
  mutate(threshold = factor(threshold, levels = c("closest 20", "closest 100", "closest 500"))) %>%
  mutate(spacetime = factor(spacetime, levels = c("other", "both", "space", "time")))

longmeansAll %>%
  ggplot(aes(x = spacetime, y = empirical_stat)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~threshold) +
  geom_text(aes(label = n), vjust=-4, size = 6)
```


Below: A plot of mean similarities of space, time, both and other words to "short" within the closest 20, 100, and 500 neighbors.

```{r}
shortmeans$threshold = "closest 500"

shortmeans20 <- shortdata %>%
  top_n(20, similarity) %>%
  get_means() %>%
  mutate(threshold = "closest 20")

shortmeans100 <- shortdata %>%
  top_n(100, similarity) %>%
  get_means() %>%
  mutate(threshold = "closest 100")

shortmeansAll <- shortmeans100 %>%
  rbind(shortmeans20) %>%
  rbind(shortmeans) %>%
  mutate(threshold = factor(threshold, levels = c("closest 20", "closest 100", "closest 500"))) %>%
  mutate(spacetime = factor(spacetime, levels = c("other", "both", "space", "time"))) 

shortmeansAll %>%
  ggplot(aes(x = spacetime, y = empirical_stat)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~threshold) +
  geom_text(aes(label = n), vjust=-4, size = 6) 

```



Below: Linear model predicting similarity to "long" by word type (space, time, etc.) in the most similar 100 words, where we're more likely to see differences. Space is the reference category. 

```{r}
longmodeldata <- longdata %>%
  top_n(100, similarity) %>%
  mutate(spacetime = relevel(spacetime, ref = 'space'))
         
longmodel <- glm(similarity ~ spacetime, family = gaussian, data = longmodeldata)

summary(longmodel)
```



Below: Linear model predicting similarity to "short" by word type (space, time, etc.) in the most similar 100 words, where we're more likely to see differences. Space is the reference category. 

```{r}
shortmodeldata <- shortdata %>%
  top_n(100, similarity) %>%
  mutate(spacetime = relevel(spacetime, ref = 'space'))
         
shortmodel <- glm(similarity ~ spacetime, family = gaussian, data = shortmodeldata)

summary(shortmodel)
```

Functions to get the counts

```{r}
getSpace <- function(data){
  return(as.numeric(data$n[data$spacetime == "space"]))
}

getTime <- function(data){
  return(as.numeric(data$n[data$spacetime == "time"]))
}

getDiff <- function(data){
  return(as.numeric(data$n[data$spacetime == "time"] - data$n[data$spacetime == "space"]))
}

```


Below: Empirical values for the number of space words, the number of time words and the difference between the two within the first 100 nearest neigbors of long.  

```{r}
empiricalValues <- longdata %>%
  top_n(100, similarity) %>%
  group_by(spacetime) %>%
  count()

observedSpace <- getSpace(empiricalValues)
observedTime <- getTime(empiricalValues)
observedDiff <- getDiff(empiricalValues)
```

Below: Empirical values for the number of space words, the number of time words and the difference between the two within the first 100 nearest neigbors of short  

```{r}
empiricalValuesShort <- shortdata %>%
  top_n(100, similarity) %>%
  group_by(spacetime) %>%
  count()

observedSpaceShort <- getSpace(empiricalValuesShort)
observedTimeShort <- getTime(empiricalValuesShort)
observedDiffShort <- getDiff(empiricalValuesShort)
```


Below: Bootstrapping (n=100) samples of number of space words, time words, and their difference from the sample of 6000 random words.

Import the random sample from the corpus with 6000 words. 

```{r message = FALSE}
null_data <- read_csv(here("neighbors/random_words.csv")) %>%
    mutate(spacetime = ifelse(is.na(spacetime), "other", spacetime)) %>%
    mutate(spacetime = as_factor(spacetime)) 
```


Generate the null distribution for number of space words, time words, and their difference. 

```{r}
set.seed(185)

sampleData <- as.data.frame(rep(0, 100))

simulate_null <- function(getX) {
    sampleData <- mutate(sampleData, spacetime = sample(null_data$spacetime, size = 100, replace=TRUE))
    sampledValues <- sampleData %>%
      group_by(spacetime, .drop = FALSE) %>%
      count() 
  return(getX(sampledValues))
}

nullSpace <- replicate(10000, simulate_null(getSpace)) 
nullTime <- replicate(10000, simulate_null(getTime))  
nullDiff <- replicate(10000, simulate_null(getDiff)) 

nullSpaceDF <- data.frame(nullSpace)
nullTimeDF <- data.frame(nullTime)
nullDiffDF <- data.frame(nullDiff)
```



```{r}
ggplot(nullTimeDF, aes(x=nullTime)) +
  geom_histogram() +
  geom_vline(xintercept = observedTime, linetype="dotted", 
                color = "blue", size=1.5) +
  geom_vline(xintercept = observedTimeShort, linetype="dotted", 
                color = "green", size=1.5)
```



```{r}
ggplot(nullSpaceDF, aes(x=nullSpace)) +
  geom_histogram() +
  geom_vline(xintercept = observedSpace, linetype="dotted", 
                color = "blue", size=1.5) +
  geom_vline(xintercept = observedSpaceShort, linetype="dotted", 
                color = "green", size=1.5)
```


```{r}
  ggplot(nullDiffDF, aes(x=nullDiff)) +
  geom_histogram() +
  geom_vline(xintercept = observedDiff, linetype="dotted", 
                color = "blue", size=1.5) +
    geom_vline(xintercept = observedDiffShort, linetype="dotted", 
                color = "green", size=1.5)
```


Below: Same analyses with the 500 observed sample and bootstrapped samples with N=500


Below: Empirical values for the number of space words, the number of time words and the difference between the two within the first 500 nearest neigbors of long.  

```{r}
empiricalValues500 <- longdata %>%
  top_n(500, similarity) %>%
  group_by(spacetime) %>%
  count()

observedSpace500 <- getSpace(empiricalValues500)
observedTime500 <- getTime(empiricalValues500)
observedDiff500 <- getDiff(empiricalValues500)
```

Below: Empirical values for the number of space words, the number of time words and the difference between the two within the first 500 nearest neigbors of short  

```{r}
empiricalValuesShort500 <- shortdata %>%
  top_n(500, similarity) %>%
  group_by(spacetime) %>%
  count()

observedSpaceShort500 <- getSpace(empiricalValuesShort500)
observedTimeShort500 <- getTime(empiricalValuesShort500)
observedDiffShort500 <- getDiff(empiricalValuesShort500)
```


Generate the null distribution for number of space words, time words, and their difference -- samples with n = 500

```{r}
set.seed(188)

sampleData <- as.data.frame(rep(0, 500))

simulate_null <- function(getX) {
    sampleData <- mutate(sampleData, spacetime = sample(null_data$spacetime, size = 500, replace=TRUE))
    sampledValues <- sampleData %>%
      group_by(spacetime, .drop = FALSE) %>%
      count() 
  return(getX(sampledValues))
}

nullSpace500 <- replicate(10000, simulate_null(getSpace)) 
nullTime500 <- replicate(10000, simulate_null(getTime))  
nullDiff500 <- replicate(10000, simulate_null(getDiff)) 

nullSpaceDF500 <- data.frame(nullSpace500)
nullTimeDF500 <- data.frame(nullTime500)
nullDiffDF500 <- data.frame(nullDiff500)
```




```{r}
ggplot(nullTimeDF500, aes(x=nullTime500)) +
  geom_histogram() +
  geom_vline(xintercept = observedTime500, linetype="dotted", 
                color = "blue", size=1.5) +
  geom_vline(xintercept = observedTimeShort500, linetype="dotted", 
                color = "green", size=1.5) +
  theme_classic() +
  labs(x = "Number of time words", y = "Count among 10000 samples") +
  #ggtitle("Observed value of number of ") + 
  geom_text(aes(x=68, label="long", y=4000), colour="blue", text=element_text(size=11))+
  geom_text(aes(x=43, label="short", y=4000), colour="green", text=element_text(size=11))

  
```


```{r}
ggplot(nullSpaceDF500, aes(x=nullSpace500)) +
  geom_histogram() +
  geom_vline(xintercept = observedSpace500, linetype="dotted", 
                color = "blue", size=1.5) +
  geom_vline(xintercept = observedSpaceShort500, linetype="dotted", 
                color = "green", size=1.5) +
  theme_classic() +
  labs(x = "Number of space words", y = "Count among 10000 samples") +
  #ggtitle("Observed value of number of ") + 
  geom_text(aes(x=24, label="long", y=4000), colour="blue", text=element_text(size=11)) +
  geom_text(aes(x=35, label="short", y=4000), colour="green", text=element_text(size=11))
```

```{r}
ggplot(nullDiffDF500, aes(x=nullDiff500)) +
  geom_histogram() +
  geom_vline(xintercept = observedDiff500, linetype="dotted", 
                color = "blue", size=1.5) +
  geom_vline(xintercept = observedDiffShort500, linetype="dotted", 
                color = "green", size=1.5) +
 theme_classic() +
  labs(x = "Number of time words minus number of space words", y = "Count among 10000 samples") +
  #ggtitle("Observed value of number of ") + 
  geom_text(aes(x=42, label="long", y=4000), colour="blue", text=element_text(size=20)) +
  geom_text(aes(x=6, label="short", y=4000), colour="green", text=element_text(size=20)) +
  theme(axis.text=element_text(size=14),
          axis.title=element_text(size=16)) 
```

Stats for long with samples N=500

For long

```{r}
mean(nullDiffDF500 >= observedDiff500)
mean(nullTimeDF500 >= observedTime500)
mean(nullSpaceDF500 >= observedSpace500)
```

For short

```{r}
mean(nullDiffDF500 >= observedDiffShort500)
mean(nullTimeDF500 >= observedTimeShort500)
mean(nullSpaceDF500 >= observedSpaceShort500)
```


