---
title: "spacetimeadultanalysis"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r load-libraries, library, message=F, results='hide', warning=FALSE}
library(knitr)
library(tidyboot)
library(tidyverse)
library(here)

theme_set(theme_classic(base_size = 16))
```

```{r message = FALSE}
# a function to get mean similarities for word types
get_means <- function(data) {
  newdata <- data %>%
    group_by(spacetime, word) %>%
    summarise_at(vars(similarity), mean) %>%
    tidyboot_mean(similarity) %>%
    mutate(spacetime = factor(spacetime, levels = c("other", "both", "space", "time"))) 
  return(newdata)
}
```

```{r message = FALSE}
# a function to get summed similarities for word types
get_sums <- function(data) {
  newdata <- data %>%
    group_by(spacetime, word) %>%
    summarise_at(vars(similarity), sum) %>%
    tidyboot(similarity, summary_function = sum,
             statistics_functions =  list("ci_lower" = ci_lower,
                                          "ci_upper" = ci_upper))
  return(newdata)
}
```


```{r message = FALSE}
read_data <- function(data) {
  cleandata <- data %>%
    mutate(spacetime = ifelse(spacetime == "O", "other",spacetime),
         spacetime = ifelse(spacetime == "S", "space",spacetime),
         spacetime = ifelse(spacetime == "T", "time",spacetime),
         spacetime = ifelse(spacetime == "B", "both",spacetime),
         spacetime = as_factor(spacetime)) %>%
    mutate(spacetime = as_factor(spacetime)) %>%
    mutate(similarity = as.numeric(similarity)) 
  return(cleandata)
}
```

```{r message = FALSE, warning=FALSE}
longdata <- read_data(read_csv(here("neighbors/cocalongneighbors.csv")))

shortdata <- read_data(read_csv(here("neighbors/cocashortneighbors.csv"))) 

widedata <- read_data(read_csv(here("neighbors/cocawideneighbors.csv"))) 

soondata <- read_data(read_csv(here("neighbors/cocasoonneighbors.csv"))) 
```



```{r}
longmeans <- get_means(longdata)
shortmeans <- get_means(shortdata)
soonmeans <- get_means(soondata)
widemeans <- get_means(widedata)

longsums <- get_sums(longdata)
shortsums <- get_sums(shortdata)
soonsums <- get_sums(soondata)
widesums <- get_sums(widedata)
```


<font size="4">Below: A histogram of word similarities to "long".</font> 


```{r}
hist(longdata$similarity)
```


<font size="4">Below: A histogram of word similarities to "short".</font> 


```{r}
hist(shortdata$similarity)
```



<font size="4">Below: A histogram of word similarities to "soon".</font> 

```{r}
hist(soondata$similarity)
```



<font size="4">Below: A histogram of word similarities to "wide".</font> 

```{r}
hist(widedata$similarity)
```





<font size="4">Below: A plot of mean similarities of space, time, both and other words to "long" within the closest 20, 100, and 500 neighbors. The numbers on the plot denote the number of words of that type in that range of neighbors.</font> 

```{r}
longmeans$threshold = "closest 500"

longmeans20 <- longdata %>%
  top_n(20, similarity) %>%
  get_means() %>%
  mutate(threshold = "closest 20")

longmeans100 <- longdata %>%
  top_n(100, similarity) %>%
  get_means() %>%
  mutate(threshold = "closest 100")

longmeansAll <- longmeans100 %>%
  rbind(longmeans20) %>%
  rbind(longmeans) %>%
  mutate(threshold = factor(threshold, levels = c("closest 20", "closest 100", "closest 500"))) %>%
  mutate(spacetime = factor(spacetime, levels = c("other", "both", "space", "time")))

longmeansAll %>%
  ggplot(aes(x = spacetime, y = empirical_stat)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~threshold) +
  geom_text(aes(label = n), vjust=-4, size = 6) +
  ggtitle("Average similarity of space, time, both and other words \nto 'long' within the closest 20, 100, and 500 neighbors") +
  labs(x = "Word type of the neighbor", y = "Average similarity to 'long'") 

```


<font size="4">Below: A plot of mean similarities of space, time, both and other words to "short" within the closest 20, 100, and 500 neighbors.</font> 

```{r}
shortmeans$threshold = "closest 500"

shortmeans20 <- shortdata %>%
  top_n(20, similarity) %>%
  get_means() %>%
  mutate(threshold = "closest 20")

shortmeans100 <- shortdata %>%
  top_n(100, similarity) %>%
  get_means() %>%
  mutate(threshold = "closest 100")

shortmeansAll <- shortmeans100 %>%
  rbind(shortmeans20) %>%
  rbind(shortmeans) %>%
  mutate(threshold = factor(threshold, levels = c("closest 20", "closest 100", "closest 500"))) %>%
  mutate(spacetime = factor(spacetime, levels = c("other", "both", "space", "time"))) 

shortmeansAll %>%
  ggplot(aes(x = spacetime, y = empirical_stat)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~threshold) +
  geom_text(aes(label = n), vjust=-4, size = 6) +
  ggtitle("Average similarity of space, time, both and other words \nto 'short' within the closest 20, 100, and 500 neighbors") +
  labs(x = "Word type of the neighbor", y = "Average similarity to 'short'") 
```


<font size="4">Below: Mean similarities of space, time, both and other words to "soon". Note: there were no pure space words for 'soon'.</font> 

```{r}
soonmeans %>%
  ggplot(aes(x = spacetime, y = empirical_stat)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  ggtitle("Average similarities to soon") +
  geom_text(aes(label = n), vjust=-4, size = 6) +
  labs(x = "Word type of the neighbor", y = "Average similarity") 

```


<font size="4">Below: Mean similarities of space, time, both and other words to "wide". Note: there were no pure time words for 'wide'.</font> 

```{r}
widemeans %>%
  ggplot(aes(x = spacetime, y = empirical_stat)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  ggtitle("Average similarities to wide") +
  geom_text(aes(label = n), vjust=-4, size = 6) +
  labs(x = "Word type of the neighbor", y = "Average similarity") 

```



<font size="4">Below: A plot of weighted sums of space, time, both and other words for "long" within the closest 20, 100, and 500 neighbors. The numbers on the plot denote the number of words of that type in that range of neighbors.</font> 

```{r}
longsums$threshold = "closest 500"

longsums20 <- longdata %>%
  top_n(20, similarity) %>%
  get_sums() %>%
  mutate(threshold = "closest 20")

longsums100 <- longdata %>%
  top_n(100, similarity) %>%
  get_sums() %>%
  mutate(threshold = "closest 100")

longsumsAll <- longsums100 %>%
  rbind(longsums20) %>%
  rbind(longsums) %>%
  mutate(threshold = factor(threshold, levels = c("closest 20", "closest 100", "closest 500"))) %>%
  mutate(spacetime = factor(spacetime, levels = c("other", "both", "space", "time")))

longsumsAll %>%
  ggplot(aes(x = spacetime, y = empirical_sum)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~threshold) +
  geom_text(aes(label = n), vjust=-4, size = 6) +
  ggtitle("Weighted sum of space, time, both and other words \nfor 'long' within the closest 20, 100, and 500 neighbors") +
  labs(x = "Word type of the neighbor", y = "Weighted sum for 'long'") 

```

<font size="4">Below: A plot of weighted sums of space, time, both and other words for "short" within the closest 20, 100, and 500 neighbors. The numbers on the plot denote the number of words of that type in that range of neighbors.</font> 


```{r}
shortsums$threshold = "closest 500"

shortsums20 <- shortdata %>%
  top_n(20, similarity) %>%
  get_sums() %>%
  mutate(threshold = "closest 20")

shortsums100 <- shortdata %>%
  top_n(100, similarity) %>%
  get_sums() %>%
  mutate(threshold = "closest 100")

shortsumsAll <- shortsums100 %>%
  rbind(shortsums20) %>%
  rbind(shortsums) %>%
  mutate(threshold = factor(threshold, levels = c("closest 20", "closest 100", "closest 500"))) %>%
  mutate(spacetime = factor(spacetime, levels = c("other", "both", "space", "time")))

shortsumsAll %>%
  ggplot(aes(x = spacetime, y = empirical_sum)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~threshold) +
  geom_text(aes(label = n), vjust=-4, size = 6) +
  ggtitle("Weighted sum of space, time, both and other words \nfor 'short' within the closest 20, 100, and 500 neighbors") +
  labs(x = "Word type of the neighbor", y = "Weighted sum for 'short'") 

```



<font size="4">Below: Summed similarities of space, time, both and other words to "soon". Note: there were no pure space words for 'soon'.</font> 

```{r}
soonsums %>%
  ggplot(aes(x = spacetime, y = empirical_sum)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  ggtitle("Summed neighbors weighted by similarity to soon") +
  geom_text(aes(label = n), vjust=-4, size = 6)
```


<font size="4">Below: Summed similarities of space, time, both and other words to "wide". Note: there were no pure time words for 'wide'.</font> 

```{r}
widesums %>%
  ggplot(aes(x = spacetime, y = empirical_sum)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  ggtitle("Summed neighbors weighted by similarity to wide") +
  geom_text(aes(label = n), vjust=-4, size = 6)
```


<font size="4">Below: Linear model predicting similarity to "long" by word type (space, time, etc.) in the most similar 100 words, where we're more likely to see differences. Space is the reference category. </font> 


```{r}
longmodeldata <- longdata %>%
  top_n(100, similarity) %>%
  mutate(spacetime = relevel(spacetime, ref = 'space'))
         
longmodel <- glm(similarity ~ spacetime, family = gaussian, data = longmodeldata)

summary(longmodel)
```



<font size="4">Below: Linear model predicting similarity to "short" by word type (space, time, etc.) in the most similar 100 words, where we're more likely to see differences. Space is the reference category. </font> 


```{r}
shortmodeldata <- shortdata %>%
  top_n(100, similarity) %>%
  mutate(spacetime = relevel(spacetime, ref = 'space'))
         
shortmodel <- glm(similarity ~ spacetime, family = gaussian, data = shortmodeldata)

summary(shortmodel)
```

<font size="4">Functions to get the counts</font> 

```{r}
getSpace <- function(data){
  return(as.numeric(data$n[data$spacetime == "space"]))
}

getTime <- function(data){
  return(as.numeric(data$n[data$spacetime == "time"]))
}

getDiff <- function(data){
  return(as.numeric(data$n[data$spacetime == "time"] - data$n[data$spacetime == "space"]))
}

```


<font size="4">Below: Empirical values for the number of space words, the number of time words and the difference between the two within the first 100 nearest neigbors of long.  </font> 

```{r}
empiricalValues <- longdata %>%
  top_n(100, similarity) %>%
  group_by(spacetime) %>%
  count()

observedSpaceLong <- getSpace(empiricalValues)
observedTimeLong <- getTime(empiricalValues)
observedDiffLong <- getDiff(empiricalValues)
```

<font size="4">Below: Empirical values for the number of space words, the number of time words and the difference between the two within the first 100 nearest neigbors of short</font>   

```{r}
empiricalValuesShort <- shortdata %>%
  top_n(100, similarity) %>%
  group_by(spacetime) %>%
  count()

observedSpaceShort <- getSpace(empiricalValuesShort)
observedTimeShort <- getTime(empiricalValuesShort)
observedDiffShort <- getDiff(empiricalValuesShort)
```


<font size="4">Below: Bootstrapping (n=100) samples of number of space words, time words, and their difference from the sample of 6000 random words.</font> 

<font size="4">Import the random sample from the corpus with 6000 words.</font> 

```{r message = FALSE}
null_data <- read_csv(here("neighbors/random_words.csv")) %>%
    mutate(spacetime = ifelse(is.na(spacetime), "other", spacetime)) %>%
    mutate(spacetime = as_factor(spacetime)) 
```


<font size="4">Generate the null distribution for number of space words, time words, and their difference. </font>

```{r}
set.seed(185)

sampleData <- as.data.frame(rep(0, 100))

simulate_null <- function(getX) {
    sampleData <- mutate(sampleData, spacetime = sample(null_data$spacetime, size = 100, replace=TRUE))
    sampledValues <- sampleData %>%
      group_by(spacetime, .drop = FALSE) %>%
      count() 
  return(getX(sampledValues))
}

nullSpace <- replicate(10000, simulate_null(getSpace)) 
nullTime <- replicate(10000, simulate_null(getTime))  
nullDiff <- replicate(10000, simulate_null(getDiff)) 

nullSpaceDF <- data.frame(nullSpace)
nullTimeDF <- data.frame(nullTime)
nullDiffDF <- data.frame(nullDiff)
```


<font size="4">Below: Same analyses with the 500 observed sample and bootstrapped samples with N=500</font>


<font size="4">Below: Empirical values for the number of space words, the number of time words and the difference between the two within the first 500 nearest neigbors of long.</font>

```{r}
empiricalValues500 <- longdata %>%
  top_n(500, similarity) %>%
  group_by(spacetime) %>%
  count()

observedSpaceLong500 <- getSpace(empiricalValues500)
observedTimeLong500 <- getTime(empiricalValues500)
observedDiffLong500 <- getDiff(empiricalValues500)
```

<font size="4">Below: Empirical values for the number of space words, the number of time words and the difference between the two within the first 500 nearest neigbors of short</font>  

```{r}
empiricalValuesShort500 <- shortdata %>%
  top_n(500, similarity) %>%
  group_by(spacetime) %>%
  count()

observedSpaceShort500 <- getSpace(empiricalValuesShort500)
observedTimeShort500 <- getTime(empiricalValuesShort500)
observedDiffShort500 <- getDiff(empiricalValuesShort500)
```


<font size="4">Generate the null distribution for number of space words, time words, and their difference -- samples with n = 500</font>  

```{r}
set.seed(188)

sampleData <- as.data.frame(rep(0, 500))

simulate_null <- function(getX) {
    sampleData <- mutate(sampleData, spacetime = sample(null_data$spacetime, size = 500, replace=TRUE))
    sampledValues <- sampleData %>%
      group_by(spacetime, .drop = FALSE) %>%
      count() 
  return(getX(sampledValues))
}

nullSpace500 <- replicate(10000, simulate_null(getSpace)) 
nullTime500 <- replicate(10000, simulate_null(getTime))  
nullDiff500 <- replicate(10000, simulate_null(getDiff)) 

nullSpaceDF500 <- data.frame(nullSpace500)
nullTimeDF500 <- data.frame(nullTime500)
nullDiffDF500 <- data.frame(nullDiff500)
```



<font size="6">TIME WORDS: among samples of 100 and 500. </font>

```{r}
ggplot(nullTimeDF, aes(x=nullTime)) +
  geom_histogram() +
  geom_vline(xintercept = observedTimeLong, linetype="dotted", 
                color = "blue", size=1.5) +
  geom_vline(xintercept = observedTimeShort, linetype="dotted", 
                color = "green", size=1.5) + 
  theme_classic() +
  geom_text(aes(x=observedTimeLong, label="observed count \nfor long", y=4000), 
            colour="blue", text=element_text(size=11)) +
  geom_text(aes(x=observedTimeShort, label="observed count \nfor short", y=4000), 
            colour="green", text=element_text(size=11)) +
  ggtitle("Null distribution for the number of time words in samples of 100") +
  labs(x = "Number of time words (out of 100)", y = "Number of samples") 
```



```{r}
ggplot(nullTimeDF500, aes(x=nullTime500)) +
  geom_histogram() +
  geom_vline(xintercept = observedTimeLong500, linetype="dotted", 
                color = "blue", size=1.5) +
  geom_vline(xintercept = observedTimeShort500, linetype="dotted", 
                color = "green", size=1.5) +
  theme_classic() +
  #ggtitle("Observed value of number of ") + 
  geom_text(aes(x=68, label="observed count \nfor long", y=4000), colour="blue", text=element_text(size=11))+
  geom_text(aes(x=43, label="observed count \nfor short", y=4000), colour="green", text=element_text(size=11)) +
  ggtitle("Null distribution for the difference in the number of time words in samples of 500") +
  labs(x = "Number of time words (out of 500)", y = "Number of samples") 
  
```


<font size="6">SPACE WORDS: among samples of 100 and 500. </font>

```{r}
ggplot(nullSpaceDF, aes(x=nullSpace)) +
  geom_histogram() +
  geom_vline(xintercept = observedSpaceLong, linetype="dotted", 
                color = "blue", size=1.5) +
  geom_vline(xintercept = observedSpaceShort, linetype="dotted", 
                color = "green", size=1.5) +
  geom_text(aes(x=observedSpaceLong, label="observed count \nfor long", y=4000), 
            colour="blue", text=element_text(size=11)) +
  geom_text(aes(x=observedSpaceShort, label="observed count \nfor short", y=4000), 
            colour="green", text=element_text(size=11)) +
  ggtitle("Null distribution for the number of space words \nin samples of 100") +
  labs(x = "Number of space words (out of 100)", y = "Number of samples") 
```


```{r}
ggplot(nullSpaceDF500, aes(x=nullSpace500)) +
  geom_histogram() +
  geom_vline(xintercept = observedSpaceLong500, linetype="dotted", 
                color = "blue", size=1.5) +
  geom_vline(xintercept = observedSpaceShort500, linetype="dotted", 
                color = "green", size=1.5) +
  theme_classic() +
  labs(x = "Number of space words", y = "Count among 10000 samples") +
  #ggtitle("Observed value of number of ") + 
  geom_text(aes(x=24, label="observed count \nfor long", y=4000), colour="blue", text=element_text(size=11)) +
  geom_text(aes(x=35, label="observed count \nfor short", y=4000), colour="green", text=element_text(size=11)) +
  ggtitle("Null distribution for the difference in the number of space words in samples of 500") +
  labs(x = "Number of space words (out of 500)", y = "Number of samples") 

```


<font size="6">TIME WORDS - SPACE WORDS: among samples of 100 and 500. </font>


```{r}
  ggplot(nullDiffDF, aes(x=nullDiff)) +
  geom_histogram() +
  geom_vline(xintercept = observedDiffLong, linetype="dotted", 
                color = "blue", size=1.5) +
    geom_vline(xintercept = observedDiffShort, linetype="dotted", 
                color = "green", size=1.5) +
  geom_text(aes(x=observedDiffLong, label="observed count \nfor long", y=4000), 
            colour="blue", text=element_text(size=11)) +
  geom_text(aes(x=observedDiffShort, label="observed count \nfor short", y=4000), 
            colour="green", text=element_text(size=11)) +
  ggtitle("Null distribution for the difference in the number of \ntime words and space words in samples of 100") +
  labs(x = "Number of time - number of space words (out of 100)", y = "Number of samples") 

```



```{r}
ggplot(nullDiffDF500, aes(x=nullDiff500)) +
  geom_histogram() +
  geom_vline(xintercept = observedDiffLong500, linetype="dotted", 
                color = "blue", size=1.5) +
  geom_vline(xintercept = observedDiffShort500, linetype="dotted", 
                color = "green", size=1.5) +
 theme_classic() +
  #labs(x = "Number of time words minus number of space words", y = "Count among 10000 samples") +
  #ggtitle("Observed value of number of ") + 
  geom_text(aes(x=42, label="observed count \nfor long", y=4000), colour="blue", text=element_text(size=20)) +
  geom_text(aes(x=6, label="observed count \nfor short", y=4000), colour="green", text=element_text(size=20)) +
  theme(axis.text=element_text(size=14),
          axis.title=element_text(size=16))  +
  ggtitle("Null distribution for the difference in the number of time words and space words \nin samples of 500") +
  labs(x = "Number of time - number of space words (out of 500)", y = "Number of samples") 

```


<font size="4">Stats with samples N=500</font>

<font size="4">For long</font>

```{r}
mean(nullDiffDF500 >= observedDiffLong500)
mean(nullTimeDF500 >= observedTimeLong500)
mean(nullSpaceDF500 >= observedSpaceLong500)
```

<font size="4">For short</font>

```{r}
mean(nullDiffDF500 >= observedDiffShort500)
mean(nullTimeDF500 >= observedTimeShort500)
mean(nullSpaceDF500 >= observedSpaceShort500)
```


